Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-07-22 BayesianCubaturePaper2018 - Rev1b.tex
fixed fourier kernel figures

\begin{abstract}
Automatic cubatures approximate multidimensional integrals to user-specified error tolerances.  For high dimensional problems, it is difficult to adaptively change the sampling pattern, but one can automatically determine the
sample size, $n$, given a reasonable, fixed sampling pattern. We take this approach here using a Bayesian perspective.  We postulate that the integrand is an instance of a Gaussian stochastic process parameterized by a constant mean and a covariance function defined by a scale parameter times a parameterized function specifying how the integrand values at two different points in the domain are related.
These hyperparameters are inferred or integrated out using integrand values via one of three techniques:  empirical Bayes, full Bayes, or generalized cross-validation. The sample size, $n$, is increased until the half-width of the credible interval for the Bayesian posterior mean is no greater than the error tolerance.

@@

We assume that the integrand, $f$, is an instance of a Gaussian stochastic process  i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$  \cite{BriEtal18a, Dia88a, OHa91a, Rit00a, RasGha03a}.  Specifically, $f$ is a real-valued random function with constant mean $m$ and covariance function $s^2C_\vtheta$:
\begin{gather*}
        m = \Ex[f(\vx)] \qquad \forall \vx \in \reals^d, \\
        \Ex\{[f(\vt) - m][f(\vx) -m]\} = s^2 C_\vtheta(\vt,\vx) \qquad \forall \vt, \vx \in \reals^d.
\end{gather*}
Here $s$ is a positive scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by the vector $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC_\vtheta^T = \mC_\vtheta,  \; \va^T \mC_\vtheta \va > 0,  \text{ where }  \mC_\vtheta = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \forall \va \ne 0, \;
 n\in \mathbb{N}, \; \text{ distinct }\vx_1, \ldots, \vx_n \in [0,1]^d.
\end{multline}
Procedures for estimating or integrating out the hyperparameters $m$, $s$, and $\vtheta$ are explained later in this section.

@@

In the theorem above, note that if the original covariance function, $C_\vtheta$, is replaced by $b C_\vtheta$ for some positive constant $b$, the cubature, $\hmu$, the estimates of $\vtheta$, and the credible interval widths, $\err_{\mathsf{x}}$ for $\mathsf{x} \in \{\MLE, \full, \GCV\}$, all remain unchanged.  The estimates of $s^2$ are multiplied by $b^{-1}$, as would be expected.

@@

	\item $\Order\bigl(N_{\opt} n^3 \bigr)$ for the matrix inversions and determinant calculations.

@@

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Not only does this variable transformation accommodate domains that are (semi-)infinite, it also tends to smooth out the integrand better than an affine transformation, which expedites the cubature.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\allowdisplaybreaks
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1),
\\
\alpha_j&(x_1,\dots,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,\dots,d,
\\
\beta_j&(x_1,\dots,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,\dots,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the univariate cumulative standard Gaussian distribution function.  Then, $\mu = \int_{[0,1]^{d'-1}} f_{\text{Genz}}(\vx) \, \dvx$. As we can see, this approach transforms a $d'$ dimensional integral into a $d=d'-1$ dimensional integral.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Transform Kernel}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We make some assumptions about the relationship between the covariance kernel and the nodes, which will be shown to hold in Section \ref{sec:shift_invariant_kernel} for rank-1 lattices and shift-invariant kernels.  Although the integrands and covariance kernels are real, it is convenient to allow related vectors and matrices to be complex.  A relevant example is the fast Fourier transform (FFT) of a real-valued vector, which is a complex-valued vector.

We introduce some notation:
\begin{align}
\nonumber
\mC = \mC_\vtheta &= \Big(C_\vtheta(\vx_i,\vx_j)\Big)_{i,j=1}^n  = (\vC_1,\dots,\vC_n) 
\\
\label{eqn:ftk_factor}
&= \frac 1n \mV \mLambda \mV^H , 
\quad \quad \mV^H = n \mV^{-1}, \\
\nonumber
\mV &= (\vv_1,\ldots,\vv_n)^T = (\vV_1,\ldots,\vV_n), \\
\nonumber
\mC^p  &= \frac 1n \mV \mLambda^{p} \mV^H, \qquad \forall p \in \integers.
\end{align}
In this and later sections, we drop the $\vtheta$ dependence of various quantities for simplicity of notation.  Here, $\mV^H$ is the Hermitian of $\mV$, $\vC_1,\dots,\vC_n$ are columns of $\mC$.  $\vV_1,\dots,\vV_n$ are columns of $\mV$ and $\vv_1,\dots,\vv_n$ are rows of $\mV$.  The normalization of $\mV$ assumed in \eqref{eqn:ftk_factor} conveniently allows the first eigenvector, $\vV_1$, to be the vector of ones in \eqref{fastcompAssumpB} below.  The columns of matrix $\mV$ are eigenvectors of $\mC$, and $\mLambda$ is a diagonal matrix of eigenvalues of $\mC$.
For any $n \times 1$ vector $\vb$, define the notation  $\widetilde{\vb} := \mV^H \vb$.

We make three assumptions that facilitate  fast computation:
\begin{subequations} \label{fastcompAssump}
	\begin{gather}
	\label{fastcompAssumpA}
	\mV \text{ may be identified analytically}, \\
	\label{fastcompAssumpB}
	\vv_1 = \vV_1 = \vone, \\
	\label{fastcompAssumpC}
	\mV^H \vb  \text{ requires only $\Order(n \log(n))$ operations } \forall \vb.
	\end{gather}
\end{subequations}
We call the transformation $\vb \mapsto \mV^H \vb$ a \emph{fast transform} and $C$ a \emph{fast transform kernel}.

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the function data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta,i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta, i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval widths simplify to
\begin{subequations}
\label{errSimple}
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\hspace{3cm} \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
\end{subequations}
In the formulas for the credible interval widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

The set of nodes used is defined by a shifted extensible integration lattice node sequence, which takes the form
\begin{equation} \label{eqn:lattice_def}
\vx_{i} = \vh \phi(i-1) + \vDelta \mod \vone, \qquad i \in \naturals.
\end{equation} 
Here, $\vh$ is a $d$-dimensional generating vector of positive integers, $\vDelta$ is some point in $[0,1)^d$, often chosen at random, and $\{\phi(i)\}_{i=0}^\infty$ is the van der Corput sequence, defined by reflecting the binary digits of the integer about the decimal point, i.e., 
\begin{align} \label{vdCDef}
\begin{array}{r|ccccccccccccc}
i & 0 & 1 & 2 & 3 & 4 &  5 & 6 & 7 & \cdots \\
i & 0_2 & 1_2 & 10_2 & 11_2 & 100_2 & 101_2 & 110_2 & 111_2  & \cdots\\
\toprule
\phi(i) & {}_2.0 &  {}_2.1 & {}_2.01 &  {}_2.11  & {}_2.001 &  {}_2.101 & {}_2.011 &  {}_2.111 & \cdots\\
\phi(i) & 0 &  0.5 &  0.25 & 0.75 &  0.125 & 0.625  &  0.375 & 0.875 & \cdots
\end{array}
\end{align}
Note that 
\begin{multline} \label{phiprop}
n\phi:\{0, \ldots, n-1 \} \to \{0, \ldots, n-1\} \\
\text{is one-to-one},
\end{multline}
assuming $n$ is a power of $2$.

@@

\subsection{Shift Invariant Kernels}
The covariance functions $C_\vtheta$ that match integration lattice node sets have the form
\begin{subequations} \label{eqn:ourkernel}
\begin{align} \label{eq:shInv}
C_\vtheta(\vt,\vx) = K_\vtheta(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift-invariant kernel} because shifting both arguments of the covariance function by the same amount leaves the value unchanged.  Here, $K_\vtheta$ is periodic and must  be of the form that ensures that $C_\vtheta$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.

@@

\subsection{The Gram Matrix as the Permutation of a Circulant Matrix}
For general shift-invariant covariance functions, the Gram matrix takes the form of a permutation of the rows and columns of a circulant matrix. By the properties of $\phi$ in \eqref{phiprop}, it follows that
\begin{equation} \label{PermMat}
    \mP = \bigl( \delta_{n\phi(i-1), j-1}  \bigr)_{i,j=1}^n
\end{equation}
is a permutation matrix, where $\delta_{\cdot,\cdot}$ is the Kronecker delta function.  Then,
\begin{align}
\nonumber
\mC_\vtheta &= \bigl ( C_\vtheta(\vx_i, \vx_j) \bigr)_{i, j = 1}^n \\
\nonumber
& = \Bigl ( K_\vtheta \bigl(\vh(\phi(i-1) - \phi(j-1) \bigr) \bmod \vone ) \Bigr)_{i, j = 1}^n \\
\nonumber
& \qquad \qquad  \text{by \eqref{eqn:lattice_def} and \eqref{eq:shInv}}  \\
\nonumber
& = \biggl( 
\sum_{i',j'=1}^n \delta_{n\phi(i-1), i'-1}  \,
K_\vtheta \bigl (\vh (i'-j')/n \bmod \vone \bigr )\\
\nonumber
& \qquad \qquad \times \delta_{j'-1,n\phi(j-1)} 
\biggr)_{i,j=1}^n \\
& = \mP \mK_{\vtheta}  \mP^T,  \label{shInvKernGramMatrix} 
\intertext{where } 
\mK_{\vtheta} &= \bigl ( K_\vtheta \bigl (\vh (i-j)/n \bmod \vone \bigr ) \bigr)_{i, j = 1}^n.
\end{align}

Because $\mK_\vtheta$ is circulant, we know the form of it's eigevector-eigenvalue decomposition:
\begin{equation} \label{Keig}
    \mK_{\vtheta} = \frac 1n \mW \mLambda_\vtheta \mW^H, \; \ \mW =  \Bigl ( \me^{2 \pi \sqrt{-1} (i-1)(j-1)/n} \Bigr)_{i,j = 1}^n.
\end{equation}
By \eqref{shInvKernGramMatrix} we then have the eigevector-eigenvalue decomposition for $\mC_{\vtheta}$ assumed in \eqref{eqn:ftk_factor}, namely
\begin{equation} \label{Clateig}
    \mC_{\vtheta} = \frac 1n \mV \mLambda_\vtheta \mV^H , \qquad \mV = \mP \mW,
\end{equation}
where the eigenvalues of $\mK_{\vtheta}$ and $\mC_{\vtheta}$ are identical.

>>>>

\begin{abstract}
Automatic cubatures approximate multidimensional integrals to user-specified error tolerances.  For high dimensional problems, it is difficult to adaptively change the sampling pattern, but one can automatically determine the
sample size, $n$, given a reasonable, fixed sampling pattern. We take this approach here using a Bayesian perspective.  We postulate that the integrand is an instance of a Gaussian stochastic process parameterized by a constant mean and a covariance kernel defined by a scale parameter times a parameterized function specifying how the integrand values at two different points in the domain are related.
These hyperparameters are inferred or integrated out using integrand values via one of three techniques:  empirical Bayes, full Bayes, or generalized cross-validation. The sample size, $n$, is increased until the half-width of the credible interval for the Bayesian posterior mean is no greater than the error tolerance.

@@

We assume that the integrand, $f$, is an instance of a Gaussian stochastic process  i.e., $f \sim \mathcal{GP}(m,s^2 C_\vtheta)$  \cite{BriEtal18a, Dia88a, OHa91a, Rit00a, RasGha03a}.  Specifically, $f$ is a real-valued random function with constant mean $m$ and covariance kernel $s^2C_\vtheta$:
\begin{gather*}
        m = \Ex[f(\vx)] \qquad \forall \vx \in \reals^d, \\
        \Ex\{[f(\vt) - m][f(\vx) -m]\} = s^2 C_\vtheta(\vt,\vx) \qquad \forall \vt, \vx \in \reals^d.
\end{gather*}
Here $s$ is a positive scale factor, and $C_\vtheta: [0,1]^d \times [0,1]^d \to \mathbb{R} $ is a symmetric, positive-definite function and parameterized by the vector $\vtheta$:
\begin{multline} \label{FJH:eq:CondPosDef}
\mC_\vtheta^T = \mC_\vtheta,  \; \va^T \mC_\vtheta \va > 0,  \text{ where }  \mC_\vtheta = \left(  C_\vtheta(\vx_i,\vx_j)  \right)_{i,j=1}^n,\\
 \forall \va \ne 0, \;
 n\in \mathbb{N}, \; \text{ distinct }\vx_1, \ldots, \vx_n \in [0,1]^d.
\end{multline}
Procedures for estimating or integrating out the hyperparameters $m$, $s$, and $\vtheta$ are explained later in this section.

@@

In the theorem above, note that if the original covariance kernel, $C_\vtheta$, is replaced by $b C_\vtheta$ for some positive constant $b$, the cubature, $\hmu$, the estimates of $\vtheta$, and the credible interval half-widths, $\err_{\mathsf{x}}$ for $\mathsf{x} \in \{\MLE, \full, \GCV\}$, all remain unchanged.  The estimates of $s^2$ are multiplied by $b^{-1}$, as would be expected.

@@

	\item $\Order\bigl(N_{\opt} n^3 \bigr)$ for the matrix inversions and determinant calculations; this cost is independent of $d$.

@@

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Not only does this variable transformation accommodate domains that are (semi-)infinite, it also tends to smooth out the integrand better, which expedites the cubature.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\allowdisplaybreaks
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1),
\\
\alpha_j&(x_1,\dots,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,\dots,d,
\\
\beta_j&(x_1,\dots,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,\dots,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the univariate cumulative standard Gaussian distribution function.  Then, $\mu = \int_{[0,1]^{d'-1}} f_{\text{Genz}}(\vx) \, \dvx$. This approach transforms a $d'$ dimensional integral into a $d=d'-1$ dimensional integral.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Transform Kernel}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We make some assumptions about the relationship between the covariance kernel and the nodes.  In Section \ref{sec:shift_invariant_kernel} these assumptions are shown to hold  for rank-1 lattices and shift-invariant kernels.  Although the integrands and covariance kernels are real, it is convenient to allow related vectors and matrices to be complex.  A relevant example is the fast Fourier transform (FFT) of a real-valued vector, which is a complex-valued vector.

We introduce some further notation:
\begin{align}
\nonumber
\mC = \mC_\vtheta &= \Big(C_\vtheta(\vx_i,\vx_j)\Big)_{i,j=1}^n  = (\vC_1,\dots,\vC_n) 
\\
\label{eqn:ftk_factor}
&= \frac 1n \mV \mLambda \mV^H , 
\quad \quad \mV^H = n \mV^{-1}, \\
\nonumber
\mV &= (\vv_1,\ldots,\vv_n)^T = (\vV_1,\ldots,\vV_n), \\
\nonumber
\mC^p  &= \frac 1n \mV \mLambda^{p} \mV^H, \qquad \forall p \in \integers.
\end{align}
In this and later sections, we drop the $\vtheta$ dependence of various quantities for simplicity of notation.  Here, $\mV^H$ is the Hermitian of $\mV$, $\vC_1,\dots,\vC_n$ are columns of $\mC$,  $\vV_1,\dots,\vV_n$ are columns of $\mV$, and $\vv_1,\dots,\vv_n$ are rows of $\mV$.  The normalization of $\mV$ assumed in \eqref{eqn:ftk_factor} conveniently allows the first eigenvector, $\vV_1$, to be the vector of ones in \eqref{fastcompAssumpB} below.  The columns of $\mV$ are eigenvectors of $\mC$, and $\mLambda$ is a diagonal matrix of eigenvalues of $\mC$.
For any $n \times 1$ vector $\vb$, define the notation  $\widetilde{\vb} := \mV^H \vb$.

We make three assumptions that facilitate  fast computation:
\begin{subequations} \label{fastcompAssump}
	\begin{gather}
	\label{fastcompAssumpA}
	\mV \text{ may be identified analytically}, \\
	\label{fastcompAssumpB}
	\vv_1 = \vV_1 = \vone, \\
	\label{fastcompAssumpC}
	\mV^H \vb  \text{ requires only $\Order(n \log(n))$ operations } \forall \vb.
	\end{gather}
\end{subequations}
We call the transformation $\vb \mapsto \mV^H \vb$ a \emph{fast transform} and $C_\vtheta$ a \emph{fast transform kernel}.

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval half-widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the function data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta,i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta, i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval half-widths simplify to
\begin{subequations}
\label{errSimple}
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\hspace{3cm} \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
\end{subequations}
In the formulas for the credible interval half-widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

We choose a set of nodes defined by a shifted extensible integration lattice node sequence, which takes the form
\begin{equation} \label{eqn:lattice_def}
\vx_{i} = \vh \phi(i-1) + \vDelta \mod \vone, \qquad i \in \naturals.
\end{equation} 
Here, $\vh$ is a $d$-dimensional generating vector of positive integers, $\vDelta$ is some point in $[0,1)^d$, often chosen at random, and $\{\phi(i)\}_{i=0}^\infty$ is the van der Corput sequence, defined by reflecting the binary digits of the integer about the decimal point, i.e., 
\begin{align} \label{vdCDef}
\begin{array}{r|ccccccccccccc}
i & 0 & 1 & 2 & 3 & 4 &  5 & 6 & 7 & \cdots \\
i & 0_2 & 1_2 & 10_2 & 11_2 & 100_2 & 101_2 & 110_2 & 111_2  & \cdots\\
\toprule
\phi(i) & {}_2.0 &  {}_2.1 & {}_2.01 &  {}_2.11  & {}_2.001 &  {}_2.101 & {}_2.011 &  {}_2.111 & \cdots\\
\phi(i) & 0 &  0.5 &  0.25 & 0.75 &  0.125 & 0.625  &  0.375 & 0.875 & \cdots
\end{array}
\end{align}
Note that 
\begin{multline} \label{phiprop}
n\phi:\{0, \ldots, n-1 \} \to \{0, \ldots, n-1\} \\
\text{is one-to-one},
\end{multline}
assuming $n$ is a power of $2$.

@@

\subsection{Shift Invariant Kernels}
The covariance kernels $C_\vtheta$ that match integration lattice node sets have the form
\begin{subequations} \label{eqn:ourkernel}
\begin{align} \label{eq:shInv}
C_\vtheta(\vt,\vx) = K_\vtheta(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift-invariant kernel} because shifting both arguments of the covariance kernel by the same amount leaves the value unchanged.  Here, $K_\vtheta$ is periodic and must  be of the form that ensures that $C_\vtheta$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.

@@

\subsection{The Gram Matrix as the Permutation of a Circulant Matrix}
For general shift-invariant covariance kernels, the Gram matrix takes the form of a permutation of the rows and columns of a circulant matrix. By the properties of $\phi$ in \eqref{phiprop}, it follows that
\begin{equation} \label{PermMat}
    \mP = \bigl( \delta_{n\phi(i-1), j-1}  \bigr)_{i,j=1}^n
\end{equation}
is a permutation matrix, where $\delta_{\cdot,\cdot}$ is the Kronecker delta function.  Then,
\begin{align}
\nonumber
\mC_\vtheta &= \bigl ( C_\vtheta(\vx_i, \vx_j) \bigr)_{i, j = 1}^n \\
\nonumber
& = \Bigl ( K_\vtheta \bigl(\vh(\phi(i-1) - \phi(j-1) \bigr) \bmod \vone ) \Bigr)_{i, j = 1}^n \\
\nonumber
& \qquad \qquad  \text{by \eqref{eqn:lattice_def} and \eqref{eq:shInv}}  \\
\nonumber
& = \biggl( 
\sum_{i',j'=1}^n \delta_{n\phi(i-1), i'-1}  \,
K_\vtheta \bigl (\vh (i'-j')/n \bmod \vone \bigr )\\
\nonumber
& \qquad \qquad \times \delta_{j'-1,n\phi(j-1)} 
\biggr)_{i,j=1}^n \\
& = \mP \mK_{\vtheta}  \mP^T,  \label{shInvKernGramMatrix} 
\intertext{where } 
\mK_{\vtheta} &= \bigl ( K_\vtheta \bigl (\vh (i-j)/n \bmod \vone \bigr ) \bigr)_{i, j = 1}^n.
\end{align}

Because $\mK_\vtheta$ is circulant, we know the form of it's eigevector-eigenvalue decomposition:
\begin{equation} \label{Keig}
    \mK_{\vtheta} = \frac 1n \mW \mLambda_\vtheta \mW^H, \; \ \mW =  \Bigl ( \me^{2 \pi \sqrt{-1} (i-1)(j-1)/n} \Bigr)_{i,j = 1}^n.
\end{equation}
By \eqref{shInvKernGramMatrix} we then have the eigevector-eigenvalue decomposition for $\mC_{\vtheta}$ assumed in \eqref{eqn:ftk_factor}, namely
\begin{equation} \label{Clateig}
    \mC_{\vtheta} = \frac 1n \mV \mLambda_\vtheta \mV^H , \qquad \mV = \mP \mW,
\end{equation}
where the eigenvalues of  $\mC_{\vtheta}$ and $\mK_{\vtheta}$ are identical.
